Standardization of ceramic shape: A case studay from the Iron Age pottery from northeastern Taiwan

### 2020-08-09 analysis/paper/paper.Rmd
added code and discussion of decoration

@@

Whether standardization suggests an emergence of specialized production depends on the number of producers, which distinguishes mechanical standardization from intentional standardization reflected by stylish and functional attributes [@Costin2001; @Costin1995]. In other words, mechanical standardization is defined by potter's skills and experiences that increased morphological uniformity normally relates to fewer producers. Population could be an indicator to estimate the potential number of producers. According to the historical records [@Nakamura1938, pp. 12], the population in Kiwulan reached a higher number in the 17th century as the European presence and declined in the Chinese period due to movement of indigenous people to the south [@Chen2007]. The changes in population size described in documents corresponds to the observation of ceramic abundance at Kiwulan. Thus, we model cv values as a function of the mass of ceramics using Poisson GLM with a link function. The model suggests that ceramic abundance strongly predicts the cv values (β = $`r beta_print`$, p = $`r pval_print`$). This indicates that a more standardized shape in the Chinese period may be influenced by a small population if ceramic abundance can reflect the population size.

In order to ensure a mechanical standardization caused by fewer producers or specialized organization, we need to investigate other factors that may contribute to intentional standardization, such as the function of ceramics. To determine if shape changes might be related to changes in the function of pots at Kiwulan, we used geochemical methods to extract and identify lipids trapped in the fabric of potsherds to identify foods that may have contributed residues absorbed into the clay [cf. @Kwak2015]. Unfortunately, we did not obtain useful results due to extremely low lipid yields, which is probably due to the very thin, dense, and low porosity fabric of Kiwulan pottery. These physical characteristics of the clay offer limited spaces to trap and protect organic molecules from microbiological degradation [cf. @Evershed2008, pp. 909].

>>>>

```{r decoration-by-phase}
pottery_deco_table <-
  pottery_data_tidy %>%
  select(Pit, Layer, `No.`, phase, `紋飾/腹片`) %>%
  mutate(`紋飾/腹片`= str_remove_all(`紋飾/腹片`, "[[:punct:]]")) %>%
  count(phase, `紋飾/腹片`, name = "count") %>%
  group_by(phase) %>%
  mutate(sum = sum(count))

pottery_deco_count <-
  pottery_deco_table %>%
  count(phase, sum, name = "decoration_types") %>%
  mutate(ratio = round(decoration_types/sum, 2))

pre_deco <- pottery_deco_count$ratio[3]
post_deco <- pottery_deco_count$ratio[2]
chinese_deco <- pottery_deco_count$ratio[1]
```

@@

Whether standardization suggests an emergence of specialized production depends on the number of producers, which distinguishes mechanical standardization from intentional standardization defined by stylish and functional attributes [@Costin2001; @Costin1995]. In other words, mechanical standardization is associated with potter's skills and experiences that increased morphological uniformity normally reflects fewer producers. Population could be an indicator to estimate the potential number of producers. According to the historical records [@Nakamura1938, pp. 12], the population in Kiwulan reached a higher number in the 17th century as the European presence and declined in the Chinese period due to movement of Indigenous people to the south [@Chen2007]. The changes in population size described in documents corresponds to the observation of ceramic abundance at Kiwulan. Thus, we model cv values as a function of the mass of ceramics using Poisson GLM with a link function. The model suggests that ceramic abundance strongly predicts the cv values (β = $`r beta_print`$, p = $`r pval_print`$). This indicates that a more standardized shape in the Chinese period may be influenced by a small population if ceramic abundance can reflect the population size.

However, intentional standardization could also contribute to shape standardization in our case. To ensure the standardization is caused by fewer producers or specialized organization, we investigate other factors that may contribute to shape standardization, such as the function and decoration of ceramics. To determine the function of pots in relation to shape changes, we used geochemical methods to extract and identify lipids trapped in the fabric of potsherds to identify foods that may have contributed residues absorbed into the clay [cf. @Kwak2015]. Unfortunately, we did not obtain useful results due to extremely low lipid yields, which is probably due to the very thin, dense, and low porosity fabric of Kiwulan pottery. These physical characteristics of the clay offer limited spaces to trap and protect organic molecules from microbiological degradation [cf. @Evershed2008, pp. 909]. Stylish attributes present a wide variation in geometric motifs from pot to pot across three phases. In general, the ceramics in the Chinese period have less types of motif according to the ratio of distinct motif to the total number for each phase (Chinese = `r chinese_deco`, post-European = `r post_deco`, pre-European = `r pre_deco`) [@WangandMarwick2020].
